---
title: "November Recap"
output: html_notebook
---

##### ***Chapter 1: Getting Started***
Install [RStudio](https://posit.co/downloads/) & [Git](https://gitforwindows.org/)

##### ***Chapter 2: Using the command line***
Get comfortable using the terminal for navigation using commands like: ls, cd, mkdir
[Reference](https://riptutorial.com/terminal/example/26023/basic-navigation-commands)

##### ***Chapter 3: Version Control***
A closer look at git/GitHub, make a repo, commit to it
Here's [My Github](https://github.com/Tomjohnsonellis)

##### ***Chapter 4: Using Markdown for Documentation***
All of the readmes for this project are made using markdown

##### ***Chapter 5: Introduction to R***

```{r}
print("Hello world!")

?function # Brings up helpful info
TRUE + FALSE # Booleans
```

##### ***Chapter 6: Functions***
```{r}
# Here's a function declaration
my_cool_function <- function(some_variable, another_variable){
  value_to_return <- some_variable + another_variable
  value_to_return # The value to be returned should be called as the last instruction
}

my_cool_function(10,5)

# Here is a function with a conditional statement
greet_user <- function(name){
  if (name == "Tom"){
    print("What is up, T-Dawg")
  } else {
    sprintf("Hello % s", name)
  }
}

greet_user("Tom")
greet_user("Joe Bloggs")
```

##### ***Chapter 7: Vectors***
A vector is a collection of data, like a list of numbers or names or dates of birth.
R is a popular data science language as the majority of functions and tools are built to work with vectorised data.
```{r}
a_vector <- c(1,2,3,4,5)
a_vector * 2
```
This approach scales up very well for working with big databases! It's both easier to follow logically and more efficient in terms of computer processing!

##### ***Chapter 8: Lists***
Lists are essentially a vector that can hold multiple data types, useful for working with related data.
```{r}
some_employee <- list(
  forename = "Joe",
  surname = "Bloggs",
  job = "CEO",
  salary = 999999,
  union_member = TRUE
)

some_employee$surname
some_employee[2]
some_employee[["surname"]]
```

##### ***Chapter 9: Understanding Data***
A more general topic of understanding where data comes from:
Sensors, surveys, record keeping
Where to find data:
Government publications (Big up the ONS), Newspapers, Scientific research, social media, online communities
Different types of data:
Nominal (apples and oranges), Ordinal (5 star rating, 4 star rating), Ratio/Continuous (Ordered with a fixed 0), Interval (Ordered but no fixed 0)

