---
title: "November Recap"
output: html_notebook
---

##### ***Chapter 1: Getting Started***
Install [RStudio](https://posit.co/downloads/) & [Git](https://gitforwindows.org/)

##### ***Chapter 2: Using the command line***
Get comfortable using the terminal for navigation using commands like: ls, cd, mkdir
[Reference](https://riptutorial.com/terminal/example/26023/basic-navigation-commands)

##### ***Chapter 3: Version Control***
A closer look at git/GitHub, make a repo, commit to it
Here's [My Github](https://github.com/Tomjohnsonellis)

##### ***Chapter 4: Using Markdown for Documentation***
All of the readmes for this project are made using markdown

##### ***Chapter 5: Introduction to R***

```{r}
print("Hello world!")

?function # Brings up helpful info
TRUE + FALSE # Booleans
```

##### ***Chapter 6: Functions***
```{r}
# Here's a function declaration
my_cool_function <- function(some_variable, another_variable){
  value_to_return <- some_variable + another_variable
  value_to_return # The value to be returned should be called as the last instruction
}

my_cool_function(10,5)

# Here is a function with a conditional statement
greet_user <- function(name){
  if (name == "Tom"){
    print("What is up, T-Dawg")
  } else {
    sprintf("Hello % s", name)
  }
}

greet_user("Tom")
greet_user("Joe Bloggs")
```

##### ***Chapter 7: Vectors***
A vector is a collection of data, like a list of numbers or names or dates of birth.
R is a popular data science language as the majority of functions and tools are built to work with vectorised data.
```{r}
a_vector <- c(1,2,3,4,5)
a_vector * 2
```
This approach scales up very well for working with big databases! It's both easier to follow logically and more efficient in terms of computer processing!

##### ***Chapter 8: Lists***
Lists are essentially a vector that can hold multiple data types, useful for working with related data.
```{r}
some_employee <- list(
  forename = "Joe",
  surname = "Bloggs",
  job = "CEO",
  salary = 999999,
  union_member = TRUE
)

some_employee$surname
some_employee[2]
some_employee[["surname"]]
```

##### ***Chapter 9: Understanding Data***
A more general topic about the world of data, covers topics such as:

* Where data comes from:
  + Sensors
  + Surveys
  + Record keeping
* Where to find data:
  + Government publications (Big up the ONS)
  + Newspapers/Journalists
  + Scientific research
  + Social media
  + Online communities
* Different types of data:
  + Nominal (apples and oranges)
  + Ordinal (5 star rating, 4 star rating)
  + Ratio/Continuous (Ordered with a fixed 0)
  + Interval (Ordered but no fixed 0)

##### ***Chapter 10: Data Frames***
Data frames are one of the most common ways to work with data, any table is a data frame, an Excel spreadsheet is a data frame.
```{r}
name <- c("Alice","Bob","Chuckie","Dylan")
height <- c(100,150,175,75)
weight <- c(50,75,100,200)

my_dataframe <- data.frame(name, height, weight)
my_dataframe

```

##### ***Chapter 11: Manipulating data with dplyr***
dplyr is a very useful and popular package for wrangling data, it has a consistent "Grammar of data manipulation"
* *Select* specific features
* *Filter* out what you don't want
* *Mutate* a data set by creating new features
* *Arrange* observations in some way
* *Summarize* data in terms of aggregates (e.g. mean, median, range, max value)
* *Join* data sets together
```{r}
library(dplyr)
my_dataframe %>%
  filter(weight < 200) %>%
  mutate("BMI" = weight / (height/100)^2) %>%
  select(name, BMI)

```

##### ***Chapter 12: Reshaping Data with tidyr***
Tidy Data:
1. Each variable is in a column
2. Each observation is a row
3. Each value is a cell
```{r}
gather() # To make your data long
spread() # To make your data wide
```

##### ***Chapter 13: Accessing Databases***
A little intro into using SQL and some info about relational databases.
Relational databases are pretty much a necessity for working with more complex data structures, requiring information to be stored in different places for different reasons. An example would be a music library, it is much more scalable to have a big database of artists and a separate big database of songs, pulling the necessary data as needed rather than having it all in one huge mess.
R can be used to work with databases, like so:
```{r}
print("Hello")
library("RSQLite") # An SQL package
library("dplyr") # Good ol' dplyr

setwd("~/GitHub/R-Practice/Exercises/chapter-13/data")
database_connection <- dbConnect(SQLite(), dbname = "path/to/database")
database_connection <- dbConnect(SQLite(), dbname = "C")
print("Hello")